---
title: "p8105_hw1_jrm2319.Rmd"
author: "Jasmin Martinez"
date: "2024-09-22"
output: html_document
---
## Homework 1 
### Problem 1
```{r setup, include=FALSE}
data("penguins", package = "palmerpenguins")
```

The data in this dataset includes the following variables: `r paste(names(penguins))`. 

The size of the dataset is `r nrow(penguins)` rows and `r ncol(penguins)` columns. 

The mean flipper lengeth is `r mean(penguins$flipper_length_mm, na.rm = TRUE)`.

Scatterplot: 
library(ggplot2) 
q1plot  <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(
    title = "Scatterplot of Flipper Length vs Bill Length",
    x = "Bill Length",
    y = "Flipper Length"
  ) +
  theme_minimal()

ggsave(filename = "scatterplot_penguins.png", plot = q1plot, width = 7, height = 5)


### Problem 2 

This is my code for the data frame and my df: 
```{r setup ,include = TRUE }
random = rnorm(10)
logical = random > 0
character = c("farm", "house", "dog", "green", "wow", "cat", "water", "hi", "bye", "welcome")
factor = factor(c("one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten"))

q2df = data.frame(
  random = random,
  logical =logical,
  character =character,
  factor = factor
)
q2df 

mean(random)
mean(logical)
mean(character)
mean(factor)
```

**Explanation**: Mean for the random and logical variable work and the means for character and factor variables do not work. Given that random is a set of random numbers that follow a normal distribution, it makes sense that a mean could be produced. A mean can also be computed from logical becuase it assigns a 1 if  random is greater than 0 and a flase is random is less than 0--however this mean is not helpful. Finally, a mean cannot be computed for chracter or factor becuase these are character variables not numeric. 

```{r w, include = TRUE }
as.numeric(character)
as.numeric(factor)
```

**Explanation**: Factor can be turned into  a numeric because it follows an ordinal syste. Character cannot be turned into a numeric becuase there is nothing that could logically make it a number in R--in the way the code is written at the moment. 